A Decision Procedure for (Co)datatypes in SMT Solvers

### 2015-02-18 tex/body.tex
cleaner
more

If $t$ is a constructor term of the form $\const{C}( u_1, \ldots, u_n )$,
then $\M( t )$ is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) }$ for some fresh $x$,
which by the induction hypothesis is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \ValC \ec{ u_1 }, \ldots, \ValC \ec{ u_n } )}$.
Call this term $t_u$.
%By Lemma~\ref{lem:interpretation-completion}\PART{2} and due to our construction of $\ValC$, we have $\ValC \ec{t}$ is a term
%of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
%where $\interp{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$ for $i \in [n]$.
Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$,
by our construction of $\ValC$ we have $\ValC \ec{t}$ is a term
of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
where for each $i \in \{1,\ldots, n\}$, % \in [n]$,
$\muvar(w_i) = {\Varec{u_i}}$.  Thus by Lemma~\ref{lem:interpretation-completion}\PART{2},
$\interp{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$.
For each $i \in \{1,\ldots, n\}$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t_u$.
By Lemma~\ref{lem:mu-norm-arg}, $\interp{t_u}{u_\iPrime} \vsim \ValC \ec{u_i}$.
Thus $\interp{t_u}{u_\iPrime} \vsim \interp{\ValC \ec{t}}{ w_i }$. % for each $i \in [n]$.
%By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
Thus, $\M( t ) \vsim t_u \vsim \ValC \ec{t}$, and
we have $\M( t ) \vsim \ValC \ec{t}$.

@@

To evaluate the decision procedure, we generated benchmark problems from existing
Isabelle formalizations using Sledgehammer.
Codatatypes being a recent addition to Isabelle
\cite{blanchette-et-al-2014-impl}, benchmarks are not abundant. We
included all the theory files from the Isabelle distribution (Distro, 879 goals)
and the \emph{Archive of Formal Proofs} (AFP, 2974 goals) \cite{klein-et-al-afp}
that define codatatypes falling within the supported fragment. We also included
two unpublished theories by Peter Gammie and Andreas Lochbihler (\gandl, 317
goals), about Bird and Stern--Brocot trees. To also exercise the support
for datatypes, the benchmarks are complemented by theories about lists
and tree data structures. The theories were selected before running the
experiments. The experimental data are publicly available.%
\footnote{\url{http://www21.in.tum.de/~blanchet/codp_data.tgz}}

@@

The raw data also reveals that the
stronger decision procedures almost completely subsume the weaker ones. For
Sledgehammer, the power of interpreted (co)datatypes is roughly comparable to
that of arithmetic \cite{blanchette-et-al-2013-smt}.

>>>>

If $t$ is a constructor term of the form $\const{C}( u_1, \ldots, u_n )$,
then $\M( t )$ is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) }$ for some fresh $x$,
which by the induction hypothesis is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \ValC \ec{ u_1 }, \ldots, \ValC \ec{ u_n } )}$.
Call this term $t_u$.
%By Lemma~\ref{lem:interpretation-completion}\PART{2} and due to our construction of $\ValC$, we have $\ValC \ec{t}$ is a term
%of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
%where $\interp{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$ for $i \in [n]$.
Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$,
by our construction of $\ValC$ we have $\ValC \ec{t}$ is a term
of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
where for each $i \in \{1,\ldots, n\}$, % \in [n]$,
$\muvar(w_i) = {\Varec{u_i}}$.  Thus by Lemma~\ref{lem:interpretation-completion}\PART{2},
$\interp{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$.
For each $i \in \{1,\ldots, n\}$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t_u$.
\begin{rep}By Lemma~\ref{lem:mu-norm-arg},\end{rep}%
\begin{conf}Clearly,\end{conf} %% TODO: Fix! ;)
$\interp{t_u}{u_\iPrime} \vsim \ValC \ec{u_i}$.
Thus $\interp{t_u}{u_\iPrime} \vsim \interp{\ValC \ec{t}}{ w_i }$. % for each $i \in [n]$.
%By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
Thus, $\M( t ) \vsim t_u \vsim \ValC \ec{t}$, and
we have $\M( t ) \vsim \ValC \ec{t}$.

@@

To evaluate the decision procedure, we generated benchmark problems from existing
Isabelle formalizations using Sledgehammer.
Codatatypes being a recent addition to Isabelle
\cite{blanchette-et-al-2014-impl}, benchmarks are not abundant. We
included all the theory files from the Isabelle distribution (Distro, 879 goals)
and the \emph{Archive of Formal Proofs} (AFP, 2974 goals) \cite{klein-et-al-afp}
that define codatatypes falling within the supported fragment. We also included
two unpublished theories by Peter Gammie and Andreas Lochbihler (\gandl, 317
goals), about Bird and Stern--Brocot trees. To exercise the support
for datatypes, the benchmarks are complemented by theories about lists
and trees. The theories were selected before running the
experiments. The experimental data are publicly available.%
\footnote{\url{http://www21.in.tum.de/~blanchet/codp_data.tgz}}

@@

The raw data also reveals that the
stronger decision procedures almost completely subsume the weaker ones. For
Sledgehammer, the practical usefulness of interpreted (co)datatypes appears
roughly comparable to that of arithmetic \cite{blanchette-et-al-2013-smt}.
